The Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament (1901--2018)

### 2019-03-23 outputs/docs/paper/paper.Rmd
Add CAP converted topics

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament (1901--2018)"
thanks: "Thank you to Chris Cochrane, Dan Simpson, Jill Sheppard, John McAndrews, John Tang, Leslie Root, Martine Mariotti, Matt Jacob, Matthew Kerby, Myles Clark, Ruth Howlett, Tianyi Wang, Tim Hatton, and Zach Ward for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. We are grateful for the many excellent comments that we received from seminar participants at the ANU SPIR, the ANU RSE, the Australian Parliamentary Library, the Max Planck Institute for Demographic Research, and the U of T Political Behavior Group. Comments and suggestions on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Monica Alexander
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: Australian National University
abstract: "A large part of politics consists of reacting to events, but not all events provoke a reaction. As such, understanding political behaviour requires examining the reaction to events and how this changes over time. We systematically analyse how discussion in the Australian Federal Parliament changes in response to two types of events: elections and changed prime ministers. To do this we first create a new dataset of what was said in the Australian Federal Parliament from 1901 through to 2018 based on available public records. We reduce the dimensionality of this dataset by using a correlated topic model, and then analyse the effect of these two types of events using a Bayesian hierarchical Dirichlet model that has several advantages over existing models. We find that: changes in prime minister tend to be associated with topic changes even when the party in power does not change; and elections that do not result in a change in prime minister are rarely associated with topic changes in the first half of our sample, but have been increasingly significant since the 1980s."
keywords: "text-as-data, Australian politics, unsupervised machine learning, Bayesian hierarchical Dirichlet model"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no

@@

@@

The percentage of stop-words each day is reasonably consistent over time (see Appendix \@ref(stopwordsgraph)). This suggests that the data are fit-for-purpose, although manual inspection does suggest there is some improvement in quality over time. We use Hansard records on a daily basis in this paper, and we pre-process our text before applying a topic model. The specific steps that we take are to: remove numbers and punctuation; change the words to lower case; and concatenate multi-word names titles and phrases, such as new south wales to new_south_wales. Then the sentences are de-constructed and each word considered individually. Not that we do not stem the words because, following @SchofieldMimno2016, we were not able to see much appreciable benefit. The resulting dataset used for analysis contains counts of words by day for 14,680 sitting days between 1901 and 2018.[^datasourcefootnote]

@@

Other work such as @BaumgartnerJones1993 and @DowdingHindmoor2010 addressed this problem by creating a standardised codebook of policy categories and sub-categories and then manually assigning text to topics as appropriate. This approach ensures the categorisation is reasonable but as it is a manual process the size of the text that can be categorised is limited.

In order to effectively categorise the topics of the entire Hansard, we chose to use topic modelling, a statistical technique which aims to extract the underlying or latent 'topics' from a collection of texts. In particular, we use a method which is similar to Latent Dirichlet Allocation (LDA), first developed by @Blei2003latent. One advantage of using the Dirichlet distribution is that it formalises the idea that there is a trade-off in what is talked about in parliament given the limited time. A multinomial approach would not have this property.

@@

An overview, and an example, of how topic modelling works is available in Appendix \@ref(LDAoverviewandexample). The underlying document generation process is discussed in Appendix \@ref(LDAdocgenprocess), and then Appendix \@ref(LDAposteriorestimation) explains how this is reversed to generate topics. Finally, Appendix \@ref(selecttopicnumber) contains a discussion of how the number of topics was chosen.

@@

One option for relating the topic distributions to events would be to use the Structural Topic Model (STM) of @RobertsStewartAiroldi2016. The distinguishing aspect of the STM is that it considers more than just a document's content when constructing topics. For example, we may believe a document's author, or the time at which it was written, or, in the case of Hansard, the prime minister or election period, may affect the topics within that document. The STM allows this additional information, or metadata, to affect the construction of topics, though influencing either topical prevalence or topical content. The assumption that there is some document generation process is the same as in LDA, it is just that this process now includes metadata.

@@

```{r topwords, echo = FALSE, results = 'asis'}

top_words %>% 
  kable("latex", 
        booktabs = T, 
        longtable = TRUE, 
        caption = "The ten words most strongly associated with each topic"
        ) %>%
  kable_styling(font_size = 8, 
                latex_options = c("repeat_header")
                )
```

@@

@@

An obvious weakness of topic models is that the number of topics needs to be specified but there is rarely a clear reason to choose some particular number. There are also more nuanced weaknesses to be aware of. For instance, the topics need to be interpreted. Some of our topics were not overly meaningful, especially on their own. Interpretation is especially difficult when the number of topics is large, but we found that a large number of topics worked best for our dataset.

@@

[^parliamentarywebsitelink]: As at 5 November, the website was available at: https://www.aph.gov.au/Parliamentary_Business/Statistics/Senate_StatsNet/General/sittingdaysyear.

@@

There are a variety of diagnostic measures that can guide the selection of the topics, but there is rarely an obvious best choice, especially at a finer level such as choosing between 60 and 65 topics. We found it useful to try a few quite different measures before settling on 80 topics. This provided a balance between being granular enough to be informative---anything less than 40 topics tended to be too broad---yet still being tractable for our analysis model in a reasonable amount of time. In addition to looking at the topics and how they changed over time, diagnostic measures that we considered include the held-out likelihood, the lower bound, residuals, exclusivity. and semantic coherence (Figures \@ref(fig:topicsdiagnostics)).[^Silgefootnote]

>>>>

---
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    fig_caption: true
    number_sections: true
    latex_engine: pdflatex
    template: svm-latex-ms.tex
title: "The Effect of Elections and Prime Ministers on Discussion in the Australian Federal Parliament (1901--2018)"
thanks: "Thank you to Chris Cochrane, Dan Simpson, Jill Sheppard, John McAndrews, John Tang, Leslie Root, Martine Mariotti, Matt Jacob, Matthew Kerby, Myles Clark, Ruth Howlett, Tianyi Wang, Tim Hatton, and Zach Ward for their helpful suggestions; and to the UC Berkeley Demography Department for the use of their computing resources. We are grateful for the many excellent comments that we received from seminar participants at the ANU SPIR, the ANU RSE, the Australian Parliamentary Library, the Max Planck Institute for Demographic Research, the UofT Political Behavior Group, and the 2019 PSA Political Methodology Conference. Comments and suggestions on the `r format(Sys.time(), '%d %B %Y')` version of this paper welcome at: rohan.alexander@anu.edu.au."
author:
- name: Rohan Alexander
  affiliation: Australian National University
- name: Monica Alexander
  affiliation: University of Toronto
abstract: "A large part of politics consists of reacting to events, but not all events provoke a reaction. As such, understanding political behaviour requires examining the reaction to events and how this changes over time. We systematically analyse how discussion in the Australian Federal Parliament changes in response to two types of events: elections and changed prime ministers. To do this we first create a new dataset of what was said in the Australian Federal Parliament from 1901 through to 2018 based on available public records. We reduce the dimensionality of this dataset using a correlated topic model, relate those topics to the Comparative Agendas Project, and then analyse the effect of these two types of events using a Bayesian hierarchical Dirichlet model that has several advantages over existing models. We find that: changes in prime minister tend to be associated with topic changes even when the party in power does not change; and elections that do not result in a change in prime minister are rarely associated with topic changes in the first half of our sample, but have been increasingly significant since the 1980s."
keywords: "text-as-data, Australian politics, unsupervised machine learning, Bayesian hierarchical Dirichlet model, Comparative Agendas Project"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
toc: FALSE
# spacing: double
bibliography: ../bibliography.bib
biblio-style: apsr
endnote: no

@@

top_words_comp <-
  read_csv2("../../../outputs/results-topic_models_and_gammas/top_words_80_comparative_agendas.csv", col_types = cols()) %>% 
  select(-`CAP number`)

@@

The percentage of stop-words each day is reasonably consistent over time (see Appendix \@ref(stopwordsgraph)). This suggests that the data are fit-for-purpose, although manual inspection does suggest there is some improvement in quality over time. We use Hansard records on a daily basis in this paper, and we pre-process our text before applying a topic model. The specific steps that we take are to: remove numbers and punctuation; change the words to lower case; and concatenate multi-word names titles and phrases, such as new south wales to new_south_wales. Then the sentences are de-constructed and each word considered individually. We do not stem the words because, following @SchofieldMimno2016, we were not able to see much appreciable benefit. The resulting dataset used for analysis contains counts of words by day for 14,680 sitting days between 1901 and 2018.[^datasourcefootnote]

@@

Other work such as @BaumgartnerJones1993 and @DowdingHindmoor2010 addressed this problem by creating a standardised codebook of policy categories and sub-categories and then manually assigning text to topics as appropriate. This approach ensures the categorisation is reasonable but as it is a manual process the size of the text that can be categorised is limited. @AshMorelliOsnabru2018 combined the best of both in the context of the New Zealand Hansard, using a supervised machine learning approach that would likly be beneficial in the Australian Hansard case as well.

In order to effectively categorise the topics of the entire Hansard, we use topic modelling, a statistical technique which aims to extract the underlying or latent 'topics' from a collection of texts. In particular, we use a method which is similar to Latent Dirichlet Allocation (LDA), first developed by @Blei2003latent. One advantage of using the Dirichlet distribution is that it formalises the idea that there is a trade-off in what is talked about in parliament given the limited time. A multinomial approach would not have this property.

@@

An overview, and an example, of how topic modelling works is available in Appendix \@ref(LDAoverviewandexample). The underlying document generation process is discussed in Appendix \@ref(LDAdocgenprocess), and then Appendix \@ref(LDAposteriorestimation) explains how this is reversed to generate topics given documents. Finally, Appendix \@ref(selecttopicnumber) contains a discussion of how the number of topics was chosen.

@@

One option for relating the topic distributions to events would be to use the Structural Topic Model (STM) of @RobertsStewartAiroldi2016. The distinguishing aspect of the STM is that it considers more than just a document's content when constructing topics. For example, we may believe a document's author, or the time at which it was written, or, in the case of our paper, the prime minister or election period, may affect the topics within that document. The STM allows this additional information, or metadata, to affect the construction of topics, though influencing either topical prevalence or topical content. The assumption that there is some document generation process is the same as in LDA, it is just that this process now includes metadata.

@@

```{r topwords, echo = FALSE, results = 'asis'}
# 
# top_words %>% 
#   kable("latex", 
#         booktabs = T, 
#         longtable = TRUE, 
#         caption = "The ten words most strongly associated with each topic"
#         ) %>%
#   kable_styling(font_size = 8, 
#                 latex_options = c("repeat_header")
#                 )
```

```{r topwordscomp, echo = FALSE, results = 'asis'}

top_words_comp %>% 
  kable("latex", 
        booktabs = T, 
        longtable = TRUE, 
        caption = "The ten words most strongly associated with each topic"
        ) %>%
  kable_styling(font_size = 7, 
                latex_options = c("repeat_header")
                )
```

@@

In the final column of Table \@ref(tab:topwords) we summarise the results of our 80-topic model using the categories of the Comparative Agendas Project (CAP), as summarised by @Bevan2017.\footnote{We use Version 1.0 of the Master Codebook, as at 31 July 2014 available at \url{http://sbevan.com/cap-master-codebook.html}.} We do this firstly, to assist other researchers in understanding our results, by trying to relate them to a a more well-known approach. We also do this because it allows us to further reduce our 80-topic model to only 21 topics.\footnote{Although the CAP Codebook topic numbering goes to 23, there are only 21 topics because there is no topic 11 or topic 22. Our 80 topics did not relate to CAP topics 21 or 23, meaning that we have 19 CAP-based categories.} It is these 19 CAP groupings to which we apply our analysis model.

@@

One weakness of topic models is that the number of topics needs to be specified but there is rarely a clear reason to choose some particular number. There are also more nuanced weaknesses to be aware of. For instance, the topics need to be interpreted. Some of our topics were not overly meaningful, especially on their own. Interpretation is especially difficult when the number of topics is large, but we found that a large number of topics worked best for our dataset.

@@

[^parliamentarywebsitelink]: As at 5 November 2018, the website was available at: https://www.aph.gov.au/Parliamentary_Business/Statistics/Senate_StatsNet/General/sittingdaysyear.

@@

There are a variety of diagnostic measures that can guide the selection of the topics, but there is rarely a clear best choice, especially at a finer level such as choosing between 60 and 65 topics. We found it useful to try a few quite different measures before settling on 80 topics. This provided a balance between being granular enough to be informative---anything less than 40 topics tended to be too broad---yet still being tractable for our analysis model in a reasonable amount of time. In addition to looking at the topics and how they changed over time, diagnostic measures that we considered include the held-out likelihood, the lower bound, residuals, exclusivity. and semantic coherence (Figures \@ref(fig:topicsdiagnostics)).[^Silgefootnote]
